---
title: "Looking at how pi changes with z"
author: "Matthew Stephens"
date: 2016-11-17
output: 
  html_document:
    css: layout/floating-toc.css
    toc: true
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

## Outline

The simplest application of ash is to z scores (so shat =1 for all observations).

In this case we can exploit the fact that for similar z the "responsibilities"
that are used in the EM algorithm will be similar to one another.

Eg suppose we evaluate the responsibilities for two different observations z1 < z2
and they are very similar. Then they will also be similar for other zs between z1 and z2.
In this way we could end up only evaluating responsibilities for a very small fraction
of all n z scores. Here we look at how similar the responsibilities are in a simple case.

```{r}
library(ashr)
set.seed(100)
n=100000
z = rnorm(n,0,2)
#now sort z so that they are in order
z = z[order(abs(z))]

res <- ash(z,1,mixcompdist="normal",outputlevel=4)
lik = res$fit_details$matrix_lik
fitted_g  = get_fitted_g(res)

# set up the initial value of pi that we typically use in ash
normalize=function(x){x/sum(x)}
pi = rep(1/n, ncomp(fitted_g))
pi[1]=1
pi = normalize(pi)

resp = apply(t(lik) * pi,2,normalize)


```

The following plots shows that the responsibilities for the first half of the 
points are very close to one another [and close to (1,0,...0)] 
```{r}
plot(resp[1,],ylim=c(0.9,1))
plot(resp[1,],ylim=c(0.999,1))
```

So how many are within 1e-5 of each responsibility?
```{r}
n_within_tol = function(i,tol=1e-5){sum(apply(abs(resp-resp[,i]),2,max)<tol)}
n_within_tol(1)
n_within_tol(25000)
n_within_tol(50000)
n_within_tol(75000)
n_within_tol(87500)
n_within_tol(93750)
```


## Session information

```{r info, echo = FALSE}
```
