---
title: "Breaking the computation time of ash into stages"
author: "Matthew Stephens"
date: 2016-11-17
output: 
  html_document:
    css: layout/floating-toc.css
    toc: true
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

## Outline

Ash has essentially three stages:
- compute the likelihood matrix (whose j,k element is p(betahat_j| component k))
- estimate g
- compute posterior quantities like E(beta_j | betahat_j, ghat)

Here we start to assess how long each step takes

First for normalmix:
```{r}
library(ashr)
set.seed(100)
z = rnorm(100000,0,2)

total_time= system.time(res <- ash(z,1,mixcompdist="normal"))
fitted_g  = get_fitted_g(res)
data = res$data

time_no_output= system.time(ash(z,1,mixcompdist="normal",outputlevel=0))
time_min_output= system.time(ash(z,1,mixcompdist="normal",outputlevel=1))
time_no_output_no_fit= system.time(ash(z,1,mixcompdist="normal",outputlevel=0,fixg=TRUE,g=fitted_g))

time_matrix = system.time(llik <- t(ashr:::log_comp_dens_conv.normalmix(fitted_g,data)))

total_time
time_matrix
time_no_output
time_min_output
time_no_output_no_fit
```

Conclusion: the initial matrix calculation is relatively negligible.
The full output takes about half the elapsed time, and a quarter of the user time.

# Same for unimix:
```{r}
total_time= system.time(res <- ash(z,1,mixcompdist="uniform"))
fitted_g  = get_fitted_g(res)
data = res$data

time_no_output= system.time(ash(z,1,mixcompdist="uniform",outputlevel=0))
time_min_output= system.time(ash(z,1,mixcompdist="uniform",outputlevel=1))

time_no_output_no_fit= system.time(ash(z,1,mixcompdist="uniform",outputlevel=0,fixg=TRUE,g=fitted_g))
time_matrix = system.time(llik <- t(ashr:::log_comp_dens_conv.unimix(fitted_g,data)))

total_time
time_matrix
time_no_output
time_min_output
time_no_output_no_fit
```

Now the full output takes more than half the elapsed time, and about half the user time.

## Session information

```{r info, echo = FALSE}
```
