---
title: "Small demo of interior-point solver"
author: "Peter Carbonetto"
date: "2017-01-24"
---

This short script demonstrates the use of the interior-point solver to
compute the solution to a quadratic program with convex objective
(*i.e.*, positive-definite Hessian) and convex, quadratic inequality
constraints. More precisely, it finds the solution to the following
optimization problem:
\begin{equation}
\begin{array}{ll}
\mbox{minimize}   & x^THx/2 + q^Tx \\
\mbox{subject to} & c_i(x) < b
\end{array}
\end{equation}
where the inequality constraints are quadratic functions:
\begin{equation}
c_i(x) = x^TP_ix/2 + r_i^Tx
\end{equation}

The quantities ${H,q,P,r,b}$ are specified below. Note that this code is not a particularly efficient way to optimize a constrained quadratic program, and 
should not be used for solving large quadratic optimization problems.

This example comes from: H. P. Schwefel (1995), *Evolution and Optimum 
Seeking.* The minimium occurs at (0,1,2,-1).

## Define the optimization problem

First, specify the quadratic objective function.

```{r define_objective}
H <- diag(c(2,2,4,2))
q <- c(-5,-5,-21,7)
```

Next, specify the quadratic inequality constraints.

```{r define_constraints}
P <- list(diag(c(4,2,2,0)),
          diag(c(2,2,2,2)),
          diag(c(2,4,2,4)))
r <- list(c(2,-1,0,-1),
          c(1,-1,1,-1),
          c(-1,0,0,-1))
b <- c(5,8,10)
```

Set the starting point.

```{r set_initial_solution}
x0 <- c(0,0,0,0)
```

## Solve the optimization problem

 Solve the optimization problem with the primal-dual interior point solver.

```{r run_ipsolver}
# TO DO: Update this code.
# data = { H q P r b };   # All the information about the quadratic program.
# z    = zeros(size(b));  # A dummy variable; does nothing.
# x    = ipsolver(x0,@(x)qprog(x,z,data,'objective'),...
#                 @(x)qprog(x,z,data,'gradient'),...
#                 @(x)qprog(x,z,data,'constraints'),...
#                 @(x,z)qprog(x,z,data,'jacobian'),...
#                 'bfgs',1e-6,100,true);
# fprintf('\nSolution:\n');
# disp(x);
```

## Session information.

```{r session_info}
print(sessionInfo())
```
