---
title: "Scaling of mixIP with n and k"
author: "Matthew Stephens"
date: 2016-11-12
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```


## Overview

Here we assess how the convex optimization in mixIP scales
with n and k.


## mixIP: Number of samples

```{r}
#define a function that calls the basic function to do mixIP in ash
mixIP = function(lik,rtol=1e-6){
  k = ncol(lik)
  prior=rep(1,k)
  prior[1]=10
  n = nrow(lik)
  init = ashr:::initpi(k,n,1)
  ashr::mixIP(lik,prior,init,ashr:::set_control_mixIP(list(rtol=rtol)))
}
```

Idea is to get a quick idea of how mixIP scales with number of samples:
```{r}
  timer_n = function(seed){
    set.seed(seed)
    z = rnorm(100000,0,2)
    #gm = exp(seq(0.01,1,length=10)) # grid of multiplier
    ll = ashr::ash(z,1,mixcompdist= "normal",outputlevel=4)
    lik = ll$fit_details$matrix_lik
    
    time = list()
    time[[1]] = system.time(mixIP(lik))
    time[[2]] = system.time(mixIP(lik[1:10000,]))
    time[[3]] = system.time(mixIP(lik[1:1000,]))
    return(time)
  }
  timer_n(1)
  timer_n(2)

```

quick conclusion: time is more than linear. We might benefit from,
say, averaging results over subsets.

## gridsizes

```{r}
 timer_k = function(seed){
    set.seed(seed)
    z = rnorm(10000,0,2)
    #gm = exp(seq(0.01,1,length=10)) # grid of multiplier
    ll = ashr::ash(z,1,mixcompdist= "normal", gridmult = 1.05, outputlevel=4)
    lik = ll$fit_details$matrix_lik
     time = list()
     k = ncol(lik)
    time[[1]] = system.time(mixIP(lik))
    time[[2]] = system.time(mixIP(lik[,c(1,sample(2:k,round(k/2),replace=F))]))
    time[[3]] = system.time(mixIP(lik[,c(1,sample(2:k,round(k/4),replace=F))]))
    time[[4]] = system.time(mixIP(lik[,c(1,sample(2:k,round(k/8),replace=F))]))
    return(time)
  }
  timer_k(1)
  timer_k(2)
```

Rough conclusion: halving k almost halves time (but not quite), so we might do well to use dense grid on small n to get plausible
grid values, and then narrow down for larger n.


## Session information

```{r session-info}
```

