---
title: "First try at fastash via Rcpp"
author: "Matthew Stephens"
date: 2016-11-22
output: 
  html_document:
    css: layout/floating-toc.css
    toc: true
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

## Overview

The file `fast_ash.cpp` contains some rcpp functions to compare.

The function `wsum_direct` simply sums the responsibilities (posterior class probabilities) directly. This is close to the approach currently used in ash.

In contrast the `wsum` method is based on a multi-resolution method that bifurcates the data
and merges rows that yield the same results within some specified tolerance.


```{r, cache=TRUE}
library(ashr)
set.seed(100)
nsamp=100000
z = rnorm(nsamp,0,2)
#now sort z so that they are in order
z = z[order(abs(z))]

res <- ash(z,1,mixcompdist="normal",outputlevel=4)
lik = res$fit_details$matrix_lik
fitted_g = get_fitted_g(res)

# set up the initial value of pi as uniform
pi = rep(1, ncomp(fitted_g))

# This is closer to the usual initial value we use in ash
# normalize=function(x){x/sum(x)}
#pi = rep(1/nsamp, ncomp(fitted_g))
#pi[1]=1
#pi = normalize(pi)


```

Here we just check the approximate methods produce similar answers to the direct method.
```{r}
Rcpp::sourceCpp('fast_ash.cpp')
wsum_direct(pi,lik,0,nsamp-1)
wsum(pi,lik,0,nsamp-1,0,0,tol=1e-3)
wsum(pi,lik,0,nsamp-1,0,0,tol=1e-5)
ws = rep(0,length(pi))
lprobsum = c(0)
add_to_wsum_direct(lprobsum, ws,pi,lik,0,nsamp-1)
ws = rep(0,length(pi))
lprobsum = c(0)
add_to_wsum(lprobsum, ws,pi,lik,0,nsamp-1,0,0,tol=1e-5)
ws

wsum_direct(pi,lik,11,2000) # just check the functions are working ok for subsets that are not the whole data 
wsum(pi,lik,11,2000,0,0,tol=1e-3)
wsum(pi,lik,11,2000,0,0,tol=1e-5)
ws = rep(0,length(pi))
lprobsum = 0
add_to_wsum(lprobsum, ws,pi,lik,11,2000,0,0,tol=1e-5)
ws
```
Here we compare the compute times. If the tolerance is too tight then
you get slower than simply directly summing. However, a weaker tolerance gives a substantial gain.
```{r}
tlik = t(lik)
system.time(wsum_direct(pi,lik,0,nsamp-1))
system.time(wsum(pi,lik,0,nsamp-1,0,0,tol=1e-3))
system.time(wsum(pi,lik,0,nsamp-1,0,0,tol=1e-5))
wsum=rep(0,length(pi))
lprobsum = 0
system.time(add_to_wsum(lprobsum,wsum,pi,lik,0,nsamp-1,0,0,tol=1e-5))
```


## Session information

```{r info, echo = FALSE}
```
